Age-related changes in plant growth and traits

### 2015-04-30 results.Rmd
more cleaning up of results

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()

source("scripts/summary_dataframes.R")
```
# file names

- leaf_counts is file with calculated leaf counts and leaf lifespan data
- SummarySppAge is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of SummarySppAge are created to omit babies or HATE for various analyses

# What causes RA to shift with age in species?
Theoretical models indicate that there need to be size (or age) related shifts in *some* trade-off to create graded RA schedules.

@@

I have both RA schedules and lots of other data across multiple ages:
  - leaf lifespa
- LMA
- accessory tissue costs

# Shifts in RA with age

Showing just a few of the RA schedules from my last lab meeting.

They show that species differ greatly in RA, that there is a general increase with age, but also that the data are noisy.

```{r, echo=FALSE,results='hide', message=FALSE}
plot_RAbyAge <- function(spp) {
  plot(RA~age, data=subset(SummaryInd,species==spp), xlab="age (years)",ylab="reproductive allocation (%)", log="x", main=labels.spp[spp],xlim=c(1,35),ylim=c(0,1))
  points(RA_mean~age, data=subset(SummarySppAge, species==spp),pch=16,cex=1.4)
}

for(n in names(labels.spp)) {
  plot_RAbyAge(n)
}
```

# Individual shoots looked decidedly "less leafy" in many older plants
This led me to wonder how leaf lifespan - or some other variable - was shifting with age

The number of leaves along a leader is overall uneffected by age, although some species show a marked decline

```{r, echo=FALSE,results='hide', message=FALSE}
plot(lvs_end_total~age,SummaryInd,log="xy", xlab="age",ylab="final leaf count along shoot",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

# Plots of individual species show that no species has a significant change either - although several show slight declines at the end

```{r, echo=FALSE,results='hide', message=FALSE}

  # fit <- lm(LL_birth ~ log10(age), data=data2)
  # abline(fit)
}

for(n in names(labels.spp)) {
  plot_LLvAge(n)
  #  mod_LMAvAge(n)
  #  summary(mod_LMAvAge(n))
}
```

# Several species show significant declines in shoot extension with age

```{r, echo=FALSE,results='hide', message=FALSE}

plot_newlengthvAge <- function(data, yvar = "new_length", xvar = "age", ...) {
	Y <- data[[yvar]]
	X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
  # points(new_length_mean ~ age, data=data, cex=1.3, pch=8)
  # points((new_length_mean+new_length_se) ~ age, data=data, cex=1.5, pch=25)
  # points((new_length_mean-new_length_se) ~ age, data=data, cex=1.5, pch=24)
}

summarise_all_fits <- function(x) {
	tmp <- lapply(x, summarise_fit) %>%
		bind_rows
	data.frame(species=names(x), tmp)
}

yvar <- "new_length"
xvar <- "age"

data <- filter(SummaryInd, age > 2.4)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(x[[yvar]] ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

for(spp in names(data)) {
  plot_newlengthvAge(data[[spp]], yvar, xvar, log="x",
  	ylim=c(0,150),  main=spp, xlab="age", ylab="leader shoot extention")
  abline(lmfits[[spp]])
}

```

# Had expected a stronger pattern either in leaf lifespan or leaf count on a shoot
Apparently it is simply that older plants have more stem - plotting total leaf weight against final weight, shows bigger plants have relatively fewer leaves
(Yes, I know the 2.4 yr olds are missing...)

```{r, echo=FALSE,results='hide', message=FALSE}
plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```
# Next steps?
As I begin to analyze RA patterns, think more specifically about the species that do show a distinct change with age in one of hte "leaf" parameters.

@@

# LMA vs Leaf lifespan
This correlation is as expected, except for the babies

```{r, echo=FALSE,results='hide', message=FALSE}

plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=17)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
```

# Growth rates increase with increasing LMA in adult plants
But there is no relationship in seedlings (age=1.4) or saplings (age=2.4)
This is not predicted by Daniel's model or Anais' review
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==1.4),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Seedlings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==1.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age==2.4),log="xy",col="pink",pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Saplings")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age==2.4), labels=species, pos=1, offset=-1.1, cex=0.9,col="blue")

plot(GrowthInv_mean ~ LMA_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="LMA",ylab="growth investment (mg per year)",main="Adults")
text(GrowthInv_mean ~ LMA_mean, data=subset(SummarySppAge,age>3), labels=species, pos=1, offset=-1.1, cex=0.6,col="blue")
legend("bottomright",col=col.age,labels.age,pch=16)
```

>>>>

```{r, echo=FALSE,results='hide', message=FALSE}
library(remake)
create_bindings()
source("scripts/summary_dataframes.R")
```
# What causes RA to shift with age in species?
Theoretical models indicate that there need to be size (or age) related shifts in *some* trade-off to create graded RA schedules.

@@

I have both RA schedules and lots of other data across multiple ages:
- leaf lifespan
- LMA
- accessory tissue costs

# Next steps?
As I begin to analyze RA patterns, think more specifically about the species that do show a distinct change with age in one of hte "leaf" parameters.

@@

### 2015-04-30 results_accessory_tissues.rmd
more cleaning up of results

#Accessory costs
Overall results:
- Accessory costs higher for bigger seeds
- A number of species show fairly consistent declines in accessory costs with various measures of "bigger", including size, age, RA. Other species show no patterns
- 9 (of 14) species decrease the proportion of reproductive investment in "pre-pollination" tissues as their RA increases. in other words, they get better returns on investment the more they invest.
- Species very different in which accessory tissues they invest the most in - which reflects the differences in basic structure (i.e. Banksia cones), but also more subtlely that some species abort lots of buds (pre-pollination aborted), while other species abort more fully formed flowers
- For most of the statistical analyses, the youngest site has been removed, since it included individuals that were harvested while flowering (and therefore never "got to" set fruit) and automatically have no "successful" reproduction. There are of course some individuals like this at other ages as well, but far fewer

#Generic functions for most plotting below
```{r, echo=FALSE,results='hide', message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

@@

#How do accessory costs shift with seed size?
Larger-seeded species have higher accessory costs. 
Also shown in other studies and suggested as a mechanism for why angiosperms were able to have such small seeds (relative to gymnosperms, with no decrease in proportional accessory costs with decreasing size)
```{r, echo=FALSE,message=FALSE}
yvar <- "PerAcc_mean"
xvar <- "seed_size"
data <- SummarySpp

@@

#Do species differ in the proportional energy allocated to propagules (vs. accessory tissues)
COER, EPMI, HEPU, LEES, PILI invest the most energy in propagules
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "species"
data <- filter(SummaryInd,Total>0)

@@

#Do species differ in the proportional energy allocated to aborted pre-pollination tissues
BOLE, EPMI, GRSP, HATE invest the most energy in aborted pre-pollination tissues
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "species"
data <- filter(SummaryInd,Total>0 & age>2)

@@

#Do individual species decrease accessory costs as they increase in size, age or RA?
First looking at size:
PerAcc is arcsine transformed, size is log transformed
COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)
GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)
PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "FinalWeight"

@@

With respect to age:
PerAcc is arcsine transformed, age is log transformed
BOLE, COER, GRSP, HATE, HEPU, PELA show a significant decrease in accessory costs with increasing age (p<0.05)
EPMI, GRBU shows a marginally significant decrease in accessory costs with increasing age (0.05<p<0.10)
PUTU shows a sig INCREASE in accessory costs with increasing age
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "age"

@@

With respect to RA:
both RA and PerAcc are arcsine transformed
BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)
BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "RA"

@@

#Next considering if the amount of energy spent on aborted buds shifts with age, size, RA
First looking at size:
PerPrePol_A is arcsine transformed, size is log transformed
GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)
PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "FinalWeight"

@@

With respect to age:
PerPrePol_A is arcsine transformed, age is log transformed
BAER, BOLE, COER, EPMI, LEES, PHPH, PILI show a significant decrease in aborted buds with increasing age (p<0.05)
GRSP, HATE show a marginally significant decrease in aborted buds with increasing age (0.05<p<0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "age"

@@

With respect to RA:
both RA and PerPrePol_A are arcsine transformed
BAER, BOLE, COER, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in aborted buds with increasing RA (p<0.05)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "RA"

@@

#Do individual species decrease total pre-pollination costs as they increase in size, age or RA?
First looking at size:
PerPrePol_All is arcsine transformed, size is log transformed
GRSP, HATE show a significant decrease in total pre-pollination costs with increasing size (p<0.05)
HEPU shows a marginally significant decrease in total pre-pollination costs with increasing size (0.05<p<0.10)

@@

With respect to age:
PerPrePol_All is arcsine transformed, age is log transformed
BAER, BOLE, COER, HEPU, LEES, PILI show a significant decrease in total pre-pollination costs with increasing age (p<0.05)
EPMI, GRBU, GRSP, HATE shows a marginally significant decrease in total pre-pollination costs with increasing age (0.05<p<0.10)
PUTU shows a sig INCREASE in total pre-pollination  costs with increasing age
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "age"

@@

With respect to RA:
both RA and PerPrePol_All are arcsine transformed
BAER, BOLE, COER, GRSP, HATE, HEPU, LEES, PEPU show a significant decrease in total pre-pollination costs with increasing RA (p<0.05)
PELA shows a marginally significant decrease in total pre-pollination costs with increasing RA (0.05<p<0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "RA"

@@

With respect to Reproductive Investment:
PerPrePol_All is arcsine transformed, ReproInv is log transformed
COER, GRBU, GRSP, HATE, PEPU, PHPH, PILI show a significant decrease in total pre-pollination costs with increasing reproductive investment (p<0.05)
LEES shows a marginally significant decrease in total pre-pollination costs with increasing reproductive investment (0.05<p<0.10)
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "ReproInv"

@@

#For each species, look at pattern of investment in different tissues across ages
Most obvious pattern is that different species invest more or less in different tissue types - across all species this is bigger pattern than shifts across age
```{r, echo=FALSE, message=FALSE}
plot_Acc_by_Spp <- function(spp) {
data2 <- filter(SummarySppAge, species==spp)
plot(PerPropagule_mean ~ age, data=data2, xlab="age", ylab="proportion tissue", cex.axis=.8, las=1,
       main=labels.spp[spp], ylim=c(0,1), xlim=c(1,35), log="x", pch=16,col="orange")
points(PerPrePol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8, col="red")
points(PerPrePol_S_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=17, col="blue")
points(PerPostPol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=18, col="black")
points(PerPackDisp_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=19, col="light green")
legend(.9,.7, pch=c(16,8,17,18,19),legend=c("Propagule","Aborted pre-poll","Successful pre-poll","Aborted post-poll","Packaging & disp"), col=c("orange","red","blue","black","light green"), cex=.8, bty="n")
}

@@

#Using models that look at all the various factors at once - but need to include species in these models and probably use some sort of a MANOVA
also issues because significance of various variables changes enormously with transformations
```{r, echo=FALSE, message=FALSE}
mod <- lm(asin(sqrt(RA))~asin(sqrt(PerPropagule))+asin(sqrt(PerPrePol_A))+ asin(sqrt(PerPrePol_S))+ asin(sqrt(PerPostPol_A))+ asin(sqrt(PerPackDisp)) + species,data=subset(SummaryInd,Total>0))

>>>>

#Accessory costs
###Overall results 
- Accessory costs higher for bigger seeds  
- A number of species show fairly consistent declines in accessory costs with various measures of "bigger", including size, age, RA. Other species show no patterns  
- 9 (of 14) species decrease the proportion of reproductive investment in "pre-pollination" tissues as their RA increases. in other words, they get better returns on investment the more they invest.  
- Species very different in which accessory tissues they invest the most in - which reflects the differences in basic structure (i.e. Banksia cones), but also more subtlely that some species abort lots of buds (pre-pollination aborted), while other species abort more fully formed flowers  
- For most of the statistical analyses, the youngest site has been removed, since it included individuals that were harvested while flowering (and therefore never "got to" set fruit) and automatically have no "successful" reproduction. There are of course some individuals like this at other ages as well, but far fewer

###Generic functions for most plotting below
```{r, echo=FALSE,results='hide', message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

@@

###How do accessory costs shift with seed size?
- Larger-seeded species have higher accessory costs. 
- Also shown in other studies and suggested as a mechanism for why angiosperms were able to have such small seeds (relative to gymnosperms, with no decrease in proportional accessory costs - with decreasing size)  
```{r, echo=FALSE,message=FALSE}
yvar <- "PerAcc_mean"
xvar <- "seed_size"
data <- SummarySpp

@@

###Do species differ in the proportional energy allocated to propagules (vs. accessory tissues)  
- COER, EPMI, HEPU, LEES, PILI invest the most energy in propagules  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "species"
data <- filter(SummaryInd,Total>0)

@@

###Do species differ in the proportional energy allocated to aborted pre-pollination tissues
- BOLE, EPMI, GRSP, HATE invest the most energy in aborted pre-pollination tissues  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "species"
data <- filter(SummaryInd,Total>0 & age>2)

@@

##Do individual species decrease accessory costs as they increase in size, age or RA?
###Change in accesory costs With size:  
- PerAcc is arcsine transformed, size is log transformed  
- COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
- GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
- PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "FinalWeight"

@@

###Change in accesory costs With age:
- PerAcc is arcsine transformed, age is log transformed  
- BOLE, COER, GRSP, HATE, HEPU, PELA show a significant decrease in accessory costs with increasing age (p<0.05)  
- EPMI, GRBU shows a marginally significant decrease in accessory costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in accessory costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "age"

@@

###Change in accesory costs With respect to RA:  
- both RA and PerAcc are arcsine transformed  
- BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
- BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "RA"

@@

##Next considering if the amount of energy spent on aborted buds shifts with age, size, RA  
###Change in aborted bud investment with size:  
- PerPrePol_A is arcsine transformed, size is log transformed  
- GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)  
- PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "FinalWeight"

@@

###Change in aborted bud investment with respect to age:  
- PerPrePol_A is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, EPMI, LEES, PHPH, PILI show a significant decrease in aborted buds with increasing age (p<0.05)  
- GRSP, HATE show a marginally significant decrease in aborted buds with increasing age (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "age"

@@

###Change in aborted bud investment with respect to RA:
- both RA and PerPrePol_A are arcsine transformed  
- BAER, BOLE, COER, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in aborted buds with increasing RA (p<0.05)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "RA"

@@

##Do individual species decrease total pre-pollination costs as they increase in size, age or RA?  
###Do total pre-pollination costs change with respect to size:  
- PerPrePol_All is arcsine transformed, size is log transformed  
- GRSP, HATE show a significant decrease in total pre-pollination costs with increasing size (p<0.05)  
- HEPU shows a marginally significant decrease in total pre-pollination costs with increasing size (0.05<p<0.10)

@@

###Do total pre-pollination costs change with respect to age:
- PerPrePol_All is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, HEPU, LEES, PILI show a significant decrease in total pre-pollination costs with increasing age (p<0.05)  
- EPMI, GRBU, GRSP, HATE shows a marginally significant decrease in total pre-pollination costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in total pre-pollination  costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "age"

@@

### Do total pre-pollination costs change with respect to RA:  
- both RA and PerPrePol_All are arcsine transformed    
- BAER, BOLE, COER, GRSP, HATE, HEPU, LEES, PEPU show a significant decrease in total pre-pollination costs with increasing RA (p<0.05)  
- PELA shows a marginally significant decrease in total pre-pollination costs with increasing RA (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "RA"

@@

### Do total pre-pollination costs change with respect to Reproductive Investment:
- PerPrePol_All is arcsine transformed, ReproInv is log transformed  
- COER, GRBU, GRSP, HATE, PEPU, PHPH, PILI show a significant decrease in total pre-pollination costs with increasing reproductive investment (p<0.05)  
- LEES shows a marginally significant decrease in total pre-pollination costs with increasing reproductive investment (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "ReproInv"

@@

###For each species, look at pattern of investment in different tissues across ages  
- Most obvious pattern is that different species invest more or less in different tissue types - across all species this is bigger pattern than shifts across age  
```{r, echo=FALSE, message=FALSE}
plot_Acc_by_Spp <- function(spp) {
data2 <- filter(SummarySppAge, species==spp)
plot(PerPropagule_mean ~ age, data=data2, xlab="age", ylab="proportion tissue", cex.axis=.8, las=1,
       main=labels.spp[spp], ylim=c(0,1), xlim=c(1,35), log="x", pch=16,col="orange")
points(PerPrePol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8, col="red")
points(PerPrePol_S_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=17, col="blue")
points(PerPostPol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=18, col="black")
points(PerPackDisp_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=19, col="light green")
legend(.9,.7, pch=c(16,8,17,18,19),legend=c("Propagule","Aborted pre-poll","Successful pre-poll","Aborted post-poll","Packaging & disp"), col=c("orange","red","blue","black","light green"), cex=.8, bty="n")
}

@@

###Considering relative effect of different accessory costs on RA
- Using models that look at all the various factors at once - but need to include species in these models and probably use some sort of a MANOVA  
- also issues because significance of various variables changes enormously with transformations  
```{r, echo=FALSE, message=FALSE}
mod <- lm(asin(sqrt(RA))~asin(sqrt(PerPropagule))+asin(sqrt(PerPrePol_A))+ asin(sqrt(PerPrePol_S))+ asin(sqrt(PerPostPol_A))+ asin(sqrt(PerPackDisp)) + species,data=subset(SummaryInd,Total>0))
